---
title: "RUM talk"
date: "12/09/2018"
output: html_document
---

Andrew Stewart, Division of Neuroscience and Experimental Psychology.

email: andrew.stewart@manchester.ac.uk

twitter: @ajstewart_lang

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First load the tidyverse and gganimate packages.

```{r, message=FALSE}
library(tidyverse)
library(gganimate)
```

Load the packages we need for our linear mixed models.

```{r, message=FALSE}
library(lme4)
library(lmerTest)
library(emmeans)
```

Load the data.

```{r, message=FALSE}
data <- read_csv("fpfp0.ixs")
```

Labelling for each condition is as follows:
C1 = Neutral condition
C2 = Negative condition
C3 = Positive condition

Relabel the Condition variable ("cond") and set as.factor()

```{r}
data$cond <- recode (data$cond, "C1"="Neutral", "C2"="Negative", "C3"="Positive")
data$cond <- as.factor(data$cond)
```

Filter so we only have the Region 3 data - this is our critical analysis region. Ignore any zero reading times - these will be tracker loss/missing data.

```{r}
data_R3 <- filter (data, reg=="R3" & DV > 0)

data_R3
```

Let's visualise the data for each of the 3 conditions.

```{r, message=FALSE}
ggplot (data_R3, aes(x=DV)) + geom_histogram() + facet_wrap(~cond)
```

Let's look at the data on a Participant by Participant basis.

```{r}
ggplot (data_R3, aes (x=cond, y=DV, colour=cond)) + geom_jitter(width=.1, alpha=.5) + 
  stat_summary(fun.y=mean, geom="point", colour="black") +
  facet_wrap(~subj) +
  scale_color_discrete(guide=FALSE) + 
  labs (y="Reading Time (msec.)", x="Condition", 
        title="Critical Region, First Pass Reading Times Facted Wrapped by Participant")
```

And via gganimate.

```{r}
ggplot (data_R3, aes (x=cond, y=DV, colour=cond)) + geom_jitter(width=.2, alpha=.8, size=3) + 
  stat_summary(fun.y=mean, geom="point", colour="black", size=5) +
  scale_color_discrete(guide=FALSE) + 
  labs (y="Reading Time (msec.)", x="Condition", 
        title="Critical Region, First Pass Reading Times \nAnimated by Participant {(closest_state)}") +
  theme(text=element_text(size=15)) +
  transition_states(subj, transition_length = 1, state_length = 1)  + 
  enter_fade() + 
  exit_fade() 
```

Let's work out for how many participants the effect went in the direction as predicted. First we're aggregating by participants.

```{r}
data_agg <- data_R3 %>% group_by(subj, cond) %>% summarise (mean=mean(DV), sd=sd(DV))
```

For how many people was the Negative condition faster than the Positive?

```{r}
sum (data_agg[data_agg$cond=="Negative",]$mean < data_agg[data_agg$cond=="Positive",]$mean)
```

For how many people was the Negative condition faster than the Neutral?

```{r}
sum (data_agg[data_agg$cond=="Negative",]$mean < data_agg[data_agg$cond=="Neutral",]$mean)
```

For how many people were the above both TRUE?

```{r}
sum (data_agg[data_agg$cond=="Negative",]$mean < data_agg[data_agg$cond=="Positive",]$mean & 
       data_agg[data_agg$cond=="Negative",]$mean < data_agg[data_agg$cond=="Neutral",]$mean)
```

Let's do the same but this time by Items.

```{r}
ggplot (data_R3, aes (x=cond, y=DV, colour=cond)) + geom_jitter(width=.1, alpha=.5) + 
  stat_summary(fun.y=mean, geom="point", colour="black") +
  facet_wrap(~item) +
  scale_color_discrete(guide=FALSE) + 
  labs (y="Reading Time (msec.)", x="Condition",
                title="Critical Region, First Pass Reading Times Facted Wrapped by Item")
```

```{r}
data_agg <- data_R3 %>% group_by(item, cond) %>% summarise (mean=mean(DV), sd=sd(DV))

sum (data_agg[data_agg$cond=="Negative",]$mean < data_agg[data_agg$cond=="Positive",]$mean)
sum (data_agg[data_agg$cond=="Negative",]$mean < data_agg[data_agg$cond=="Neutral",]$mean)
```

Let's build a raincloud plot.

```{r, warning=FALSE, message=FALSE}
library(RColorBrewer)
library(plyr) #note, need to detach this after this plot as clashes with aspects of dplyr
source("https://gist.githubusercontent.com/ajstewartlang/6c4cd8ab9e0c27747424acdfb3b4cff6/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

```{r, warning=FALSE}
raincloud_theme = theme(
  text = element_text(size = 12),
  axis.title.x = element_text(size = 12),
  axis.title.y = element_text(size = 12),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(angle = 45, vjust = 0.5),
  legend.title=element_text(size=12),
  legend.text=element_text(size=12),
  legend.position = "right",
  #plot.title = element_text(lineheight=.8, size = 12),
  panel.border = element_blank(),
  panel.grid.minor = element_blank(),
  panel.grid.major = element_blank(),
  axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
  axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'))

lb <- function(x) mean(x) - sd(x)
ub <- function(x) mean(x) + sd(x)

sumld <- ddply(data_R3, ~cond, summarise, mean = mean(DV), median = median(DV), lower = lb(DV), upper = ub(DV))

ggplot(data = data_R3, aes(y = DV, x = cond, fill = cond)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0), alpha = .8, trim=FALSE) +
  geom_point(aes(y = DV, color = cond), position = position_jitter(width = .15), size = .5, alpha = 0.8) +
  geom_boxplot(width = .1,  outlier.shape = NA, alpha = 0.5) +
  #expand_limits(x = 3) +
  guides(fill = FALSE) +
  guides(color = FALSE) +
  scale_color_brewer(palette = "Accent") +
  scale_fill_brewer(palette = "Accent") +
  coord_flip() +
  theme_bw() +
  raincloud_theme +
  labs(y="Reading Time (msec.)", x="Condition", title="Critical Region, First Pass Reading Times")

detach("package:plyr", unload=TRUE) #need to detach as some clashes with dplyr
```

Build our first linear mixed model with maximal random effects structure.

```{r}
model.null <- lmer(DV ~ (1+cond|subj) + (1+cond|item), data_R3, REML=TRUE)
model <- lmer(DV ~ cond + (1+cond|subj) + (1+cond|item), data_R3, REML=TRUE)
```

Check the experimental model differs from the Null model.

```{r}
anova(model.null, model)
```

Generate a summary of our model parameters plus run some pairwise comparisons.

```{r}
summary(model)
emmeans(model, pairwise~cond, adjust="none")
```

Check we're happy with the model residuals.

```{r}
qqnorm(residuals(model))
```

I can live with these...

Now let's look at the subsequent region of text (Region 4). Exclude missing data/tracker loss points where the DV==0.

```{r}
data_R4 <- filter (data, reg=="R4" & DV > 0)
```

Need to load the plyr package again.

```{r, warning=FALSE, message=FALSE}
library(plyr)
```

Another Raincloud plot.

```{r, warning=FALSE}
sumld <- ddply(data_R4, ~cond, summarise, mean = mean(DV), median = median(DV), lower = lb(DV), upper = ub(DV))

ggplot(data = data_R4, aes(y = DV, x = cond, fill = cond)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0), alpha = .8, trim=FALSE) +
  geom_point(aes(y = DV, color = cond), position = position_jitter(width = .15), size = .5, alpha = 0.8) +
  geom_boxplot(width = .1,  outlier.shape = NA, alpha = 0.5) +
  #expand_limits(x = 3) +
  guides(fill = FALSE) +
  guides(color = FALSE) +
  scale_color_brewer(palette = "Accent") +
  scale_fill_brewer(palette = "Accent") +
  coord_flip() +
  theme_bw() +
  raincloud_theme +
  labs(y="Reading Time (msec.)", x="Condition", title="Post-Critical Region, First Pass Reading Times")

detach("package:plyr", unload=TRUE) #need to detach as some clashes with dplyr
```

Build the linear mixed model for Region 4 as we did for Region 3.

```{r}
model.null <- lmer(DV ~ (1|subj) + (1|item), data_R4, REML=TRUE)
model <- lmer(DV ~ cond + (1|subj) + (1|item), data_R4, REML=TRUE)

anova (model.null, model)

summary (model)
emmeans (model, pairwise~cond, adjust="none")
```

Looking at the first pass data animated by Region.

```{r}
data1 <- filter(data, reg!="R0" & DV > 0)

ggplot (data1, aes (x=cond, y=DV, colour=cond)) + geom_jitter(width=.2, alpha=.8, size=3) + 
  geom_boxplot(width = .5,  outlier.shape = NA, alpha = 0.3, colour="black") +
  scale_color_discrete(guide=FALSE) + 
  labs (y="Reading Time (msec.)", x="Condition", 
        title="First Pass Reading Times Animated by Region. \nThe Critical Region is R3. \nCurrent Region is {(closest_state)}") +
  theme(text=element_text(size=15)) +
  coord_flip() +
  transition_states(reg, transition_length = 1, state_length = 1)  + 
  enter_fade() + 
  exit_fade() 
```


Read in the Regressions Out data.

```{r}
FPRO <- read_csv("fproro0.ixs")
```

Recode as before.

```{r}
data_R3 <- filter (FPRO, reg=="R3" & cond != "C0")
data_R3$cond <- recode (data_R3$cond, "C1"="Neutral", "C2"="Negative", "C3"="Positive")
data_R3$cond <- as.factor (data_R3$cond)
```

Let's generate some summary data.

```{r}
data_summ <- data_R3 %>% group_by(cond) %>% summarise(mean=mean(DV), sd=sd(DV))
```

Build a bar chart.

```{r}
ggplot (data_summ, aes (x=cond, y=mean, fill=cond)) + geom_col() + 
  geom_text(aes (label= format(round(mean, 0))), nudge_y = 1) +
  scale_fill_discrete (guide=FALSE) + labs (y="Regressions (%)", x="Condition", 
                                            title="Regressions Out of Critical Region")
```

Recode the DV as binomial.

```{r}
data_R3$DV <- recode (data_R3$DV, "100"=1 ,"0"=0)
```

Build the null and experimental glmm given sampling from the binomial distribution.

```{r}
model.null <- glmer (DV ~ (1|subj) + (1|item), data_R3, family=binomial)
model <- glmer (DV ~ cond + (1|subj) + (1|item), data_R3, family=binomial)
anova (model.null, model)
summary (model)
```

Now look at Regressions Out of the post-critical region (Region 4).

```{r}
data_R4 <- filter (FPRO, reg=="R4" & cond != "C0")
data_R4$cond <- recode (data_R4$cond, "C1"="Neutral", "C2"="Negative", "C3"="Positive")
data_R4$cond <- as.factor (data_R4$cond)

data_summ <- data_R4 %>% group_by(cond) %>% summarise(mean=mean(DV), sd=sd(DV))
```

Build a bar chart.

```{r}
ggplot (data_summ, aes (x=cond, y=mean, fill=cond)) + geom_col() + 
  geom_text(aes (label= format(round(mean, 0))), nudge_y = 1) +
  scale_fill_discrete (guide=FALSE) + labs (y="Regressions (%)", x="Condition",
                                            title="Regressions Out of Post-Critical Region")
```

Recode DV as binomial.

```{r}
data_R4$DV <- recode (data_R4$DV, "100"=1 ,"0"=0)
```

Build the glmms.

```{r}
model.null <- glmer (DV ~ (1|subj) + (1|item), data_R4, family=binomial)
model <- glmer (DV ~ cond + (1|subj) + (1|item), data_R4, family=binomial)

anova (model.null, model)
summary (model)
```

As model is significant, run some pairwise comparisions to determine where the effect is. Ask for the output on the response scale.

```{r}
emmeans (model, pairwise~cond, adjust="none", type="response")
```

Animate % Regressions Out by Region.

```{r}
data <- filter (FPRO, cond != "C0" & reg!="R0" & reg!="R9" & reg!="R8" & reg!="R7" & reg!="R6")

data$cond <- recode (data$cond, "C1"="Neutral", "C2"="Negative", "C3"="Positive")
data$cond <- as.factor (data$cond)

data_summ <- data %>% group_by(reg, subj, cond) %>% summarise(mean=mean(DV))

ggplot (data_summ, aes (x=cond, y=mean, colour=cond)) + geom_point(alpha=.8, size=3) + 
  scale_color_discrete(guide=FALSE) + 
  geom_boxplot(width = .5,  outlier.shape = NA, alpha = 0.3, colour="black") +
  labs (y="Regressions Out (%)", x="Condition", 
        title="% of Regressions Out Animated by Region. \nThe Critical Region is R3. \nCurrent Region is {(closest_state)}") +
  theme(text=element_text(size=15)) +
  coord_flip() +
  transition_states(reg, transition_length = 1, state_length = 1)  + 
  enter_fade() + 
  exit_fade() 
```


